---
title: "Regression results"
author: "Ivy"
date: "2025-01-16"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(fs)
library(here)
library(readxl)
library(writexl)
library(tidyverse)
library(lme4)
library(broom)
library(emmeans)
library(plotly)
library(ggplot2)
library(superb)
library(pwr)
```

# Read in data

Three models:

1. LSTM
2. CNN
3. residual linear network

Two training conditions:

1. low-passed filter --> full-frequency
2. high-passed filter --> full-frequency
3. full-frequency

Three test data sets:

1. training set: h/l filter
2. target testing set: h/l filter --> h+l filter
3. full testing set: h+l filter

```{r}
# model 3 * train_filter_cond 2 * test_type 3 * preepoch_cond 17 * run 20 * epoch 121 = 740520 entries
accstage <- 
  read_csv(here("/Users/ivy/Desktop/Modeling/LowFrequency/data/accstage-final.csv")) %>%
  rename("model" = model_cond) %>% 
  pivot_longer(cols = c("train_acc", "target_valid_acc", "full_valid_acc"),
               names_to = "test_type", values_to = "acc") %>% 
  mutate(train_filter_cond = if_else(train_filter_cond == "lf", "low", "high"),
         test_type = case_when(test_type == "train_acc" ~ "train",
                               test_type == "target_valid_acc" ~ "target testing",
                               test_type == "full_valid_acc" ~ "full testing")) %>% 
  mutate(preepoch_cond = factor(preepoch_cond,
                                levels = c("0", "1", "2", "3", "4", "5", "10", "15", "20", "25", "30", "35", "40", "45", "50", "55", "60")),
         model = factor(model, levels = c("lstm", "cnn", "resnet")),
         train_filter_cond = factor(train_filter_cond, levels = c("low", "high")),
         test_type = factor(test_type, levels = c("train", "target testing", "full testing"))) %>%
  select(model, train_filter_cond, test_type, preepoch_cond, run, epoch, acc)
accstage
```

# Establish convergence {.tabset}

## Stablizing slope at the end of Stage I
```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 = 240 entries
accstage_slope_60 <- 
  accstage %>% 
  # focus on validation sets in Stage I of condition 60
  filter(preepoch_cond == 60,
         epoch <= 60 - 1,
         test_type != "train") %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # document the increase in accuracy of the last five epochs
  mutate(acc_inc = acc - lag(acc),
         acc_inc_2 = lag(acc_inc),
         acc_inc_3 = lag(acc_inc_2),
         acc_inc_4 = lag(acc_inc_3),
         acc_inc_5 = lag(acc_inc_4)) %>% 
  filter(epoch == 59)
accstage_slope_60
```

```{r}
# set Exp1 stablizing slope
slope_1 = 0.03
```

## Stablizing slope at the end of Stage II
```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 = 240 entries
accstage_slope_0 <- 
  accstage %>% 
  # focus on validation sets in Stage II of condition 0
  filter(preepoch_cond == 0,
         test_type != "train") %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # document the increase in accuracy of the last five epochs
  mutate(acc_inc = acc - lag(acc),
         acc_inc_2 = lag(acc_inc),
         acc_inc_3 = lag(acc_inc_2),
         acc_inc_4 = lag(acc_inc_3),
         acc_inc_5 = lag(acc_inc_4)) %>% 
  filter(epoch == 119)
accstage_slope_0
```

```{r}
# set Exp2 stablizing slope
slope_2 = 0.01
```

# EXP1

Two training stages:

1. Stage I: 15 epochs
2. Stage II: 15 epochs

Three check points:

1. CP1: 12-14 epochs
2. CP2: 17-19 epochs
3. CP3: 27-29 epochs

```{r}
# set variables
first_stage_end = 15 # before birth
second_stage_mid = 20 # short period after birth
second_stage_end = 30 # when learning completes
average_epoch = 3 # calculate average of 3 consecutive epochs
```

```{r}
# model 3 * train_filter_cond 3 * test_type 3 * run 20 * epoch 31 = 16740 entries
accstage_15 <- 
  accstage %>% 
  filter(preepoch_cond == first_stage_end | 
           (preepoch_cond == 0 & train_filter_cond == "low"),
         epoch <= second_stage_end - 1) %>% 
  mutate(train_filter_cond = if_else(preepoch_cond == 0, "full", train_filter_cond),
         train_filter_cond = factor(as.factor(train_filter_cond),
                                    levels = c("low", "high", "full"))) %>% 
  select(-preepoch_cond)
accstage_15
```

## Learning trajectory
```{r}
# model 3 (grid_y) * train_filter_cond 3 (grid_x) * test_type 3 (line) * epoch 31 (x) = 837 entries
dataset_sum <- 
  accstage_15 %>% 
  group_by(model, train_filter_cond, test_type, epoch) %>% 
  summarise(mean = mean(acc),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
dataset_sum
```

```{r}
dataset_learning_plot <- 
  ggplot(data = dataset_sum) +
  geom_line(aes(x = epoch, y = mean, color = test_type)) +
  geom_ribbon(aes(x = epoch, ymin = ci_min, ymax = ci_max, fill = test_type), alpha = 0.25) +
  facet_grid(model ~ train_filter_cond) +
  scale_x_continuous(breaks = seq(0, 60, by = 2)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2),
                     limits = c(0, 1)) +
  labs(x = "epochs", y = "accuracy rates", color = "dataset", fill = "dataset")
dataset_learning_plot
```

## Check point 1 {.tabset}

```{r}
# model 3 * train_filter_cond 2 * test_type 3 * run 20 * epoch 3 = 1080 entries
accstage_1214 <- 
  accstage_15 %>% 
  filter(train_filter_cond != "full",
         epoch >= first_stage_end - average_epoch & epoch <= first_stage_end - 1)
accstage_1214
```

```{r}
# model 3 * train_filter_cond 2 * test_type 3 = 18 entries
sum_1214 <- 
  accstage_1214 %>% 
  group_by(train_filter_cond, test_type, model) %>% 
  summarise(mean = mean(acc),
            show_mean = paste0(format(round(100*mean, digits = 2), nsmall = 2), "%"),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se) %>% 
  mutate(epoch = case_when(train_filter_cond == "low" ~ 12.5,
                           train_filter_cond == "high" ~ 13.5))
sum_1214
```

### model
```{r}
mdl_1214 <- 
  lm(acc ~ train_filter_cond * test_type * model, data = accstage_1214)
mdl_1214 %>% summary()
```

```{r}
f2 = (0.9813 / (1 - 0.9813)) * (17 / 1062)
pwr.f2.test(u = 17, v = 1062, f2 = f2, sig.level = 0.05)
```

```{r}
pairwise_1214 <- 
  emmeans(mdl_1214, pairwise ~ train_filter_cond | test_type * model, adjust = "mvt")
pairwise_1214
```

```{r}
pairwise_output_1214 <- 
  pairwise_1214$contrasts %>% 
  tidy() %>% 
  select(test_type:contrast, estimate:p.value) %>% 
  mutate(across(estimate:statistic, ~ round(.x, 2)),
         p.value = scales::pvalue(p.value)) %>% 
  mutate(significance = case_when(p.value <= 0.001 ~ "***",
                                  p.value <= 0.01 ~ "**",
                                  p.value <= 0.05 ~ "*",
                                  p.value <= 0.1 ~ ".",
                                  p.value > 0.1 ~ "N.S."),
         model = factor(model, levels = c("lstm", "cnn", "resnet")),
         test_type = factor(test_type, levels = c("train", "target testing", "full testing"))) %>% 
  left_join(sum_1214 %>% 
              filter(train_filter_cond == "low") %>% 
              select(test_type:mean))
```

### plot
```{r}
plot_1214 <- 
  ggplot(data = sum_1214, aes(x = train_filter_cond, y = mean)) +
  geom_bar(stat = "identity", aes(fill = train_filter_cond)) +
  geom_errorbar(aes(x = train_filter_cond, ymax = ci_max, ymin = ci_min),
                width = 0.5, color = "grey50") +
  geom_text(y = sum_1214$mean + 0.05, label = sum_1214$show_mean,
            color = "grey50", size = 3) +
  # showSignificance(c(1,2), 0.5, 0, "",
  #                  segmentParams = list(color = "grey50"),
  #                  textParams = list(color = "grey50")) +
  # geom_text(x = 1.5, y = 0.55, label = sum_1214$significance,
  #           color = "grey50", size = 3) +
  facet_grid(model ~ test_type) +
  scale_y_continuous(breaks = seq(0, 0.6, by = 0.2),
                     limits = c(0, 0.6)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = "condition", y = "accuracy rates", fill = "condition") +
  theme(legend.position = "None")
plot_1214
```

## Check point 2 {.tabset}

```{r}
# model 3 * train_filter_cond 2 * test_type 3 * run 20 * epoch 3 = 1080 entries
accstage_1719 <- 
  accstage_15 %>% 
  filter(train_filter_cond != "full",
         epoch >= second_stage_mid - average_epoch & epoch <= second_stage_mid - 1)
accstage_1719
```

```{r}
# model 3 * train_filter_cond 2 * test_type 3 = 18 entries
sum_1719 <- 
  accstage_1719 %>% 
  group_by(train_filter_cond, test_type, model) %>% 
  summarise(mean = mean(acc),
            show_mean = paste0(format(round(100*mean, digits = 2), nsmall = 2), "%"),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se) %>% 
  mutate(epoch = case_when(train_filter_cond == "low" ~ 17.5,
                           train_filter_cond == "high" ~ 18.5))
sum_1719
```

### model
```{r}
mdl_1719 <- 
  lmer(acc ~ train_filter_cond * test_type * model + (1 | run), data = accstage_1719)
mdl_1719 %>% summary()
```

```{r}
f2 = (0.9325 / (1 - 0.9325)) * (17 / 1062)
pwr.f2.test(u = 17, v = 1062, f2 = f2, sig.level = 0.05)
```

```{r}
pairwise_1719 <- 
  emmeans(mdl_1719, pairwise ~ train_filter_cond | test_type * model, adjust = "mvt")
pairwise_1719
```

```{r}
pairwise_output_1719 <- 
  pairwise_1719$contrasts %>% 
  tidy() %>% 
  select(test_type:contrast, estimate:p.value) %>% 
  mutate(across(estimate:statistic, ~ round(.x, 2)),
         p.value = scales::pvalue(p.value)) %>% 
  mutate(significance = case_when(p.value <= 0.001 ~ "***",
                                  p.value <= 0.01 ~ "**",
                                  p.value <= 0.05 ~ "*",
                                  p.value <= 0.1 ~ ".",
                                  p.value > 0.1 ~ "N.S."),
         model = factor(model, levels = c("lstm", "cnn", "resnet")),
         test_type = factor(test_type, levels = c("train", "target testing", "full testing"))) %>% 
  left_join(sum_1719 %>% 
              filter(train_filter_cond == "low") %>% 
              select(test_type:mean))
```

### plot
```{r}
plot_1719 <- 
  ggplot(data = sum_1719, aes(x = train_filter_cond, y = mean)) +
  geom_bar(stat = "identity", aes(fill = train_filter_cond)) +
  geom_errorbar(aes(x = train_filter_cond, ymax = ci_max, ymin = ci_min),
                width = 0.5, color = "grey50") +
  geom_text(y = sum_1719$mean - 0.08, label = sum_1719$show_mean,
            color = "grey50", size = 3) +
  # showSignificance(c(1,2), 0.7, 0, "",
  #                  segmentParams = list(color = "grey50"),
  #                  textParams = list(color = "grey50")) +
  # geom_text(x = 1.5, y = 0.75, label = sum_1719$significance,
  #           color = "grey50", size = 3) +
  facet_grid(model ~ test_type) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2)) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = "condition", y = "accuracy rates", fill = "condition") +
  theme(legend.position = "None")
plot_1719
```

## Check point 3 {.tabset}

```{r}
# model 3 * train_filter_cond 2 * test_type 3 * run 20 * epoch 3 = 1080 entries
accstage_2729 <- 
  accstage_15 %>% 
  filter(train_filter_cond != "full",
         epoch >= second_stage_end - average_epoch & epoch <= second_stage_end - 1)
accstage_2729
```

```{r}
# model 3 * train_filter_cond 2 * test_type 3 = 18 entries
sum_2729 <- 
  accstage_2729 %>% 
  group_by(train_filter_cond, test_type, model) %>% 
  summarise(mean = mean(acc),
            show_mean = paste0(format(round(100*mean, digits = 2), nsmall = 2), "%"),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se) %>% 
  mutate(epoch = case_when(train_filter_cond == "low" ~ 27.5,
                           train_filter_cond == "high" ~ 28.5))
sum_2729
```

### model
```{r}
mdl_2729 <- 
  lmer(acc ~ train_filter_cond * test_type * model + (1 | run), data = accstage_2729)
mdl_2729 %>% summary()
```

```{r}
f2 = (0.9836 / (1 - 0.9836)) * (17 / 1062)
pwr.f2.test(u = 17, v = 1062, f2 = f2, sig.level = 0.05)
```

```{r}
pairwise_2729 <- 
  emmeans(mdl_2729, pairwise ~ train_filter_cond | test_type * model, adjust = "mvt")
pairwise_2729
```

```{r}
pairwise_output_2729 <- 
  pairwise_2729$contrasts %>% 
  tidy() %>% 
  select(test_type:contrast, estimate:p.value) %>% 
  mutate(across(estimate:statistic, ~ round(.x, 2)),
         p.value = scales::pvalue(p.value)) %>% 
  mutate(significance = case_when(p.value <= 0.001 ~ "***",
                                  p.value <= 0.01 ~ "**",
                                  p.value <= 0.05 ~ "*",
                                  p.value <= 0.1 ~ ".",
                                  p.value > 0.1 ~ "N.S."),
         model = factor(model, levels = c("lstm", "cnn", "resnet")),
         test_type = factor(test_type, levels = c("train", "target testing", "full testing"))) %>% 
  left_join(sum_2729 %>% 
              filter(train_filter_cond == "low") %>% 
              select(test_type:mean))
```

### plot
```{r}
plot_2729 <- 
  ggplot(data = sum_2729, aes(x = train_filter_cond, y = mean)) +
  geom_bar(stat = "identity", aes(fill = train_filter_cond)) +
  geom_errorbar(aes(x = train_filter_cond, ymax = ci_max, ymin = ci_min),
                width = 0.5, color = "grey50") +
  geom_text(y = sum_2729$mean - 0.08, label = sum_2729$show_mean,
            color = "grey50", size = 3) +
  # showSignificance(c(1,2), 0.9, 0, "",
  #                  segmentParams = list(color = "grey50"),
  #                  textParams = list(color = "grey50")) +
  # geom_text(x = 1.5, y = 0.95, label = sum_2729$significance,
  #           color = "grey50", size = 3) +
  facet_grid(model ~ test_type) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2)) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = "condition", y = "accuracy rates", fill = "condition") +
  theme(legend.position = "None")
plot_2729
```

## Three check points combined
```{r}
# model 3 (grid_y) * train_filter_cond 2 (line) * test_type 3 (grid_x) * epoch 31 (x) = 558 entries
h_l_sum <- 
  accstage_15 %>% 
  filter(train_filter_cond != "full") %>% 
  group_by(model, train_filter_cond, test_type, epoch) %>% 
  summarise(mean = mean(acc),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
h_l_sum
```

```{r}
three_stages_plot <- 
  ggplot() +
  # learning curve
  geom_line(data = h_l_sum, 
            aes(x = epoch, y = mean, color = train_filter_cond)) +
  geom_ribbon(data = h_l_sum, 
              aes(x = epoch, ymin = ci_min, ymax = ci_max, fill = train_filter_cond), alpha = 0.25) +
  # end of first stage
  geom_bar(data = sum_1214, 
           aes(x = epoch, y = mean, fill = train_filter_cond),
           stat = "identity", position = "identity", width = 2.5) + 
  geom_errorbar(data = sum_1214, 
                aes(x = epoch, ymax = ci_max, ymin = ci_min),
                width = 2, linewidth = 0.3, color = "grey30") +
  geom_text(data = pairwise_output_1214,
            aes(x = 13, y = mean + 0.05, label = significance),
            color = "grey30", size = 8, size.unit = "pt") +
  # mid of second stage
  geom_bar(data = sum_1719, 
           aes(x = epoch, y = mean, fill = train_filter_cond),
           stat = "identity", position = "identity", width = 2.5) + 
  geom_errorbar(data = sum_1719, 
                aes(x = epoch, ymax = ci_max, ymin = ci_min),
                width = 2, linewidth = 0.3, color = "grey30") +
  geom_text(data = pairwise_output_1719,
            aes(x = 18, y = mean + 0.05, label = significance),
            color = "grey30", size = 8, size.unit = "pt") +
  # end of second stage
  geom_bar(data = sum_2729,
           aes(x = epoch, y = mean, fill = train_filter_cond),
           stat = "identity", position = "identity", width = 2.5) +
  geom_errorbar(data = sum_2729,
                aes(x = epoch, ymax = ci_max, ymin = ci_min),
                width = 2, linewidth = 0.3, color = "grey30") +
  geom_text(data = pairwise_output_2729,
            aes(x = 28, y = mean + 0.05, label = significance),
            color = "grey30", size = 8, size.unit = "pt") +
  # stages
  geom_text(data = data.frame(x = c(7.5, 22.5),
                              y = c(0.95, 0.95),
                              text = c("Stage I", "Stage II")),
            aes(x, y, label = text),
            color = "grey30", size = 8, size.unit = "pt") +
  geom_vline(xintercept = 15, linetype = 3, color = "grey30") +
  facet_grid(model ~ test_type) +
  scale_x_continuous(breaks = seq(0, 30, by = 5)) +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0, 1, by = 0.2),
                     limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  theme_gray(base_size = 10) +
  labs(x = "epochs", y = "accuracy rates", color = "condition", fill = "condition")
three_stages_plot
```


## Confirm convergence {.tabset}

### End of Stage I
```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 * epoch 16 = 3840 entries
accstage_slope_15_stage_1 <- 
  accstage_15 %>% 
  # focus on only validation sets in stage I of non-full conditions
  filter(train_filter_cond != "full",
         test_type != "train",
         epoch <= first_stage_end - 1) %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # document the increase in accuracy in previous five epochs
  mutate(acc_inc = acc - lag(acc),
         acc_inc_2 = lag(acc_inc),
         acc_inc_3 = lag(acc_inc_2),
         acc_inc_4 = lag(acc_inc_3),
         acc_inc_5 = lag(acc_inc_4))
accstage_slope_15_stage_1
```

```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 = 240 entries
accstage_stable_15_stage_1 <- 
  accstage_slope_15_stage_1 %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # include all epochs stablizing at 0.03 slope for more than 3 epochs
  filter(-slope_1 <= acc_inc, acc_inc <= slope_1,
         -slope_1 <= acc_inc_2, acc_inc_2 <= slope_1,
         -slope_1 <= acc_inc_3, acc_inc_3 <= slope_1) %>% 
  # select the first of stablizing epochs
  filter(epoch == first(epoch))
accstage_stable_15_stage_1
```

```{r}
# model 3 * train_filter_cond 2 * test_type 2 = 12 entries
stable_sum_15_stage_1 <- 
  accstage_stable_15_stage_1 %>% 
  group_by(model, train_filter_cond, test_type) %>% 
  summarise(mean = mean(epoch),
            n = length(epoch),
            sd = sd(epoch),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
stable_sum_15_stage_1
```

### End of Stage II
```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 * epoch 15 = 3600 entries
accstage_slope_15_stage_2 <- 
  accstage_15 %>% 
  # focus on only validation sets in stage I of non-full conditions
  filter(train_filter_cond != "full",
         test_type != "train",
         first_stage_end <= epoch,
         epoch <= second_stage_end - 1) %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # document the increase in accuracy in previous five epochs
  mutate(acc_inc = acc - lag(acc),
         acc_inc_2 = lag(acc_inc),
         acc_inc_3 = lag(acc_inc_2),
         acc_inc_4 = lag(acc_inc_3),
         acc_inc_5 = lag(acc_inc_4))
accstage_slope_15_stage_2
```

```{r}
# model 3 * train_filter_cond 2 * test_type 2 * run 20 = 240 entries
accstage_stable_15_stage_2 <- 
  accstage_slope_15_stage_2 %>% 
  group_by(model, train_filter_cond, test_type, run) %>% 
  # include all epochs stablizing at 0.02 slope for more than 3 epochs
  filter(-slope_1 <= acc_inc, acc_inc <= slope_1,
         -slope_1 <= acc_inc_2, acc_inc_2 <= slope_1,
         -slope_1 <= acc_inc_3, acc_inc_3 <= slope_1) %>% 
  # select the first of stablizing epochs
  filter(epoch == first(epoch))
accstage_stable_15_stage_2
```

```{r}
# model 3 * train_filter_cond 2 * test_type 2 = 12 entries
stable_sum_15_stage_2 <- 
  accstage_stable_15_stage_2 %>% 
  group_by(model, train_filter_cond, test_type) %>% 
  summarise(mean = mean(epoch),
            n = length(epoch),
            sd = sd(epoch),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
stable_sum_15_stage_2
```

## Confirm learning

```{r}
# model 3 * train_filter_cond 3 * run 20 * epoch 4 = 720 entries
baseline <- 
  accstage_15 %>% 
  filter(test_type == "full testing",
         epoch == -1 | (epoch >= second_stage_end - average_epoch & epoch <= second_stage_end - 1)) %>% 
  mutate(train_filter_cond = if_else(epoch == -1, "baseline", train_filter_cond))
baseline
```

```{r}
# model 3 * train_filter_cond 4 = 12 entries
baseline_sum <- 
  baseline %>% 
  group_by(model, train_filter_cond) %>% 
  summarise(mean = mean(acc),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
baseline_sum
```

```{r}
baseline_mdl <- 
  lmer(acc ~ train_filter_cond * model + (1 | run), data = baseline)
baseline_mdl %>% summary()
```

```{r}
baseline_pairwise <- 
  emmeans(baseline_mdl, pairwise ~ train_filter_cond | model, adjust = "mvt")
baseline_pairwise
```

## Output tables
```{r}
table_1 <-
  baseline_pairwise$contrasts %>% 
  tidy() %>% 
  select(model, contrast, estimate:adj.p.value) %>% 
  mutate(across(estimate:statistic, ~ round(.x, 3)),
         adj.p.value = scales::pvalue(adj.p.value))
```

```{r}
table_2 <-
  pairwise_1214$contrasts %>% tidy() %>% mutate(check_point = "CP1") %>% 
  full_join(pairwise_1719$contrasts %>% tidy() %>% mutate(check_point = "CP2")) %>% 
  full_join(pairwise_2729$contrasts %>% tidy() %>% mutate(check_point = "CP3")) %>% 
  select(check_point, model, test_type, contrast, estimate:p.value) %>% 
  rename(data_set = "test_type") %>% 
  mutate(across(estimate:statistic, ~ round(.x, 3)),
         p.value = scales::pvalue(p.value))
```

```{r}
write_xlsx(table_1, path = "plots/table_1.xlsx")
write_xlsx(table_2, path = "plots/table_2.xlsx")
```

## Output plots
```{r}
ggsave("plots/three stages.pdf", three_stages_plot, width = 190, height = 160, units = "mm",dpi = 300)
```

# EXP2 different pre-trainning epochs

## Learning trajectory
```{r}
# model 3 (grid_y) * train_filter_cond 2 (grid_x) * test_type 3 (grid_x) * preepoch_cond 17 (line) * epoch 121 (x) = 37026 entries
preepoch_sum <- 
  accstage %>% 
  group_by(model, train_filter_cond, test_type, preepoch_cond, epoch) %>% 
  summarise(mean = mean(acc),
            show_mean = paste0(format(round(100*mean, digits = 2), nsmall = 2), "%"),
            n = length(acc),
            sd = sd(acc),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
preepoch_sum
```

```{r}
preepoch_learning_plot <- 
  ggplot(data = preepoch_sum) +
  geom_line(aes(x = epoch, y = mean, color = preepoch_cond)) +
  geom_ribbon(aes(x = epoch, ymin = ci_min, ymax = ci_max, fill = preepoch_cond), alpha = 0.25) +
  facet_grid(model ~ interaction(test_type, train_filter_cond)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2)) +
  labs(x = "epochs", y = "accuracy rates", color = "conditions", fill = "conditions")
preepoch_learning_plot
```

## Learning speed {.tabset}

### plot
```{r}
# model 3 * test_type 2 * preepoch_cond 17 * run 20 * epoch 121 = 246840 entries
accstage_slope <- 
  accstage %>% 
  # focus on only low conditions
  filter(train_filter_cond == "low") %>% 
  # focus on only validation sets
  filter(test_type != "train") %>% 
  group_by(model, train_filter_cond, test_type, preepoch_cond, run) %>% 
  # document the increase in accuracy in previous ten epochs
  mutate(acc_inc = acc - lag(acc),
         acc_inc_2 = lag(acc_inc),
         acc_inc_3 = lag(acc_inc_2),
         acc_inc_4 = lag(acc_inc_3),
         acc_inc_5 = lag(acc_inc_4),
         post_epoch = as.numeric(epoch) - as.numeric(as.character(preepoch_cond)))
accstage_slope
```

```{r}
# model 3 * test_type 2 * preepoch_cond 17 * run 20 = 2040 entries
accstage_stable <- 
  accstage_slope %>% 
  group_by(model, train_filter_cond, test_type, preepoch_cond, run) %>% 
  # filter out all stage 1 training epochs
  filter(post_epoch >= 0) %>% 
  # include all epochs stablizing at 0.01 slope for more than 3 epochs
  filter(-slope_2 <= acc_inc, acc_inc <= slope_2,
         -slope_2 <= acc_inc_2, acc_inc_2 <= slope_2,
         -slope_2 <= acc_inc_3, acc_inc_3 <= slope_2) %>% 
  # select the first of stablizing epochs
  filter(epoch == first(epoch))
accstage_stable
```

```{r}
# model 3 * test_type 2 * preepoch_cond 17 = 102 entries
stable_sum <- 
  accstage_stable %>% 
  group_by(model, train_filter_cond, test_type, preepoch_cond) %>% 
  summarise(mean = mean(post_epoch),
            n = length(post_epoch),
            sd = sd(post_epoch),
            se = sd / sqrt(n),
            t = qt((0.95)/2 + 0.5, n-1),
            ci_min = mean - t * se,
            ci_max = mean + t * se)
stable_sum
```

```{r}
preepoch_speed_table <- 
  stable_sum %>% 
  select(model, train_filter_cond, test_type, preepoch_cond, mean) %>% 
  mutate(mean = as.integer(mean)) %>% 
  pivot_wider(names_from = preepoch_cond, values_from = mean)
preepoch_speed_table
```

```{r}
preepoch_speed_plot <- 
  ggplot(data = stable_sum, aes(x = preepoch_cond, y = mean)) +
  geom_point(aes(color = preepoch_cond)) +
  geom_path(group = 1) +
  geom_errorbar(aes(ymax = ci_max, ymin = ci_min, color = preepoch_cond)) +
  facet_grid(model ~ interaction(test_type, train_filter_cond)) +
  scale_y_continuous(breaks = seq(0, 30, by = 5)) +
  labs(x = "condition", y = "average number of epochs to reach convergence", color = "condition")
preepoch_speed_plot
```

### model
```{r}
# focus preepoch_cond
focus_cond = 15

# get all nine grids
groups <- unique(paste(accstage_stable$model, accstage_stable$test_type, sep = "_"))
groups

# apply function to all nine grids in low filter condition
lapply(groups, function(group) {
  lm(post_epoch ~ preepoch_cond,
     data = accstage_stable %>% 
       filter(model == str_split(group, "_")[[1]][1],
              train_filter_cond == "low",
              test_type == str_split(group, "_")[[1]][2],
              preepoch_cond == 0 | preepoch_cond == focus_cond)) %>% 
    summary()
})
```
